Computing solution space properties of combinatorial optimization problems via generic tensor networks

### 2022-09-03 paper/paper.tex
update notation

\section{Tensor networks}\label{sec:tn}
A tensor network is a multi-linear map composed of a collection of labelled tensors $\mathcal{T}$ and an indicator for the output tensor. It is formally defined as follows.
\begin{definition}[Tensor Network~\cite{Cirac2021, Orus2014}]
    A tensor network is a triple of $\mathcal{N} = (\Sigma, \Tensors, O)$.
    $\Sigma$ is an alphabet of labels.
    $\Tensors$ is a set of labelled tensors as the inputs, each $T_{\boldsymbol{\sigma}} \in \mathcal{T}$ is labeled by a vector of labels $\boldsymbol{\sigma} \in \Sigma^{r(T)}$, where $r(T)$ the rank of $T$.
    $O$ is a labbeled tensor as the output.
\end{definition}
The operation to evaluate a tensor network is called \textit{contraction}, which is defined as a summation of tensor element products over the labels not appearing in the output tensor.
For example, the matrix multiplication is a special tensor network
\begin{equation}
\mathcal{N}_{\rm matmul} = \left(\{i,j,k\}, \{A_{ij}, B_{jk}\}, C_{ik}\right),
\end{equation}
where $A, B$ and $C$ are matrices (two-dimensional tensors), among which $C$ is the output tensor, $(i,k),(i,j)$ and $(i,k)$ are labels associated with them. The contraction is defined as $C_{ik} = \sum_j A_{ij}B_{jk}$, where the subscripts are for tensor indexing, and the tensor dimensions with a same label must have the same size.
%can be viewed as a generalization of binary matrix multiplication to nary tensor contraction.
%Generalized Einstein's notation is often used to represent a tensor network, where we use a label in the subscript to represent a degree of freedom.
%One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% Einstein's notation is often used to represent a tensor network, e.g.\ it represents the matrix multiplication between two matrices $A$ and $B$ as $C_{ik} = A_{ij}B_{jk}$,
% where we use a label in the subscript to represent a degree of freedom.
% One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice, either both in input tensors (which will be summed over) or one in a input tensor and another in the output tensor.
% Hence a tensor network can be represented as a open simple graph,
% where a tensor is mapped to a vertex, a label shared by two input tensors is mapped to a normal edge and a label appears in the output tensor is mapped to an open edge.
% In this work, we adopt a more general form of tensor networks that do not restrict how many times a label can appear in tensors. 
The graphical representation of a tensor network is an open (meaning an edge can connect to external vertices) hypergraph, where an input tensor is mapped to a vertex and a label is mapped to a hyperedge that can connect an arbitrary number of vertices, while the labels appearing in the output tensor are open.
Our notation is a minor generalization of the standard tensor network notation used in physics as we do not restrict the number of times a label can appear in the tensors. 
While this generalized form is equivalent in representation power, it can have smaller contraction complexity as will be illustrated in \Cref{sec:tensorbad}.
% because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors, a high dimensional equivalence of the identity matrix.
%However, introducing $\delta$ tensors may significantly increase the complexity of the tensor contraction. We illustrate this subtle point in \Cref{sec:tensorbad}.

\begin{example}
\blue{maybe remove this example?}
\begin{equation}
\begin{split}
    \Sigma &= \{i,j,k,l,m\},\\
    \mathcal{T} &= \{A_{jkm}, B_{mil}, V_{jm}\},\\
    O &= C_{ijk},\\
    \end{split}
\end{equation}
is a tensor network that can be evaluated as $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$.
Its hypergraph representation is shown below, where we use different colors to represent different hyperedges.

@@

\section{Tensor network representation of independent sets} \label{sec:tnmap}
This section describes the reduction of the independent set problem to a tensor network contraction problem.
An alternative interpretation, perhaps more accessible to physicists, can be found in \Cref{sec:energymodel}, where we introduce the reduction from the energy model of hardcore lattice gases~\cite{Dyre2016, Fernandes2007}.
Let $G=(V,E)$ be an undirected graph with each vertex $v\in V$ being associated with a weight $w_v$.
An independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, we have $(u, v) \not\in E$.
To the independent set problem on $G$ can be encoded as a tensor network $\mathcal{N}_{\rm IS}(G)$
\begin{equation}\label{eq:mistensornetwork}
\begin{split}
    \Sigma &= \{s_v \mid v \in V\},\\
    \mathcal{T} &= \{W^{(v)}_{v} \mid v\in V\} \cup \{B^{(u, v)}_{uv} \mid (u, v) \in E\},\\
    O &= P,
    \end{split}
\end{equation}
where for each vertex $v$, we define a parameterized rank-one tensor associated with it as
\begin{equation}
    W^{(v)} = \left(\begin{matrix}
        1 \\
        x_v^{w_v}
    \end{matrix}\right),\label{eq:vertextensor}
\end{equation}
and for each edge $(u, v) \in E$, we define a matrix $B$ indexed by $s_u$ and $s_v$ as
\begin{equation}
    \qquad \quad 
       B^{(u, v)} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right). \label{eq:edgetensor}
\end{equation}
We map each vertex $v\in V$ to a label $s_v \in \{0, 1\}$ of dimension $2$, where we use $0$ ($1$) to denote a vertex is absent (present) in $I$.
%We use the superscripts to denote which vertex or edge a tensor is associated to.
These labels can be used as subscripts of tensors to index tensor elements, e.g.\ $W^{(v)}_0=1$ is the first element associated with $s_v=0$ and $W^{(v)}_1=x_v^{w_v}$ is the second element associated with $s_v=1$, where $x_v$ is a variable associated with vertex $v$.
$P$ is the output tensor of rank $0$, i.e. a scalar.
The independence constraint is encoded in the edge tensors, where we use $B_{11} = 0$ to denote two vertices connected by an edge cannot be both in the independent set. 
The contraction of this tensor network is
\begin{equation}\label{eq:idp}
    P(G) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{v\in V} W^{(v)}_{s_v} \prod\limits_{(u,v) \in E} B^{(u,v)}_{s_u s_v},
\end{equation}
where the summation runs over all $2^{|V|}$ vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$. 
% It is easy to see that an edge tensor element $B_{s_us_v}$ encodes the independent set constraint. It zeros out the product whenever both $u$ and $v$ are in the independent set.
A vertex tensor element $W^{(v)}_{s_v}$ contributes a multiplicative factor $x_v^{w_v}$ whenever $v$ is in the set.

@@

\begin{example}\label{eg:tensorcontraction}
\blue{maybe remove this example too?}
In this example, we map a 5-vertex graph (left) to a tensor network (right) and show how the contraction order reduces the time and space complexities.

@@

Let $x_i = x$ and $w_i = 1$, \Cref{eq:idp} corresponds to the independence polynomial:
\begin{equation}\label{eq:idpdef}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$, $\alpha(G) \equiv \max_{I}\sum_{v\in I}w_v$ is the size of a maximum independent set and it is also called the independence number.
An independence polynomial is a useful graph characteristic related to, for example, the partition functions~\cite{Lee1952,Yang1952} and Euler characteristics of the independence complex~\cite{Bousquet2008, Levit2009}.
%Its connection to \Cref{eq:idp} can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
%and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. The summation counts the number of independent sets of size $k$. 
By assigning a real number to $x$, one can evaluate this independence polynomial for this specific value directly using tensor network contraction.
For example, the total number of independent sets can be evaluated as $I(G, 1)$.
However, instead of evaluating this polynomial for a certain value, we are more interested in knowing the coefficients of this polynomial, because this quantity tells us the counting of independent sets at different sizes.
%first elevate the tensor elements $0$s and $1$s in tensors $W(x)$ and $B$ from the standard number types to the additive identity, $\mymathbb{0}$, and multiplicative identity, $\mymathbb{1}$, of a commutative semiring as discussed in \Cref{sec:generic}.
%\purple{I'd
%say that you're about to describe this semiring, otherwise it seems like you're going to
%use the general definition of the semiring rather than a specific type of semiring}\blue{from $1$ to $\mymathbb{1}$ and $0$ to $\mymathbb{0}$ applies for a general semiring, I haven't change this part, let's discuss to make sure I understand your point.}\purple{Perhaps start the sentence with: Our semiring elements are a polynomial type...}
To this end, let us create a polynomial number data type by representing a polynomial $a_0 + a_1 x + \ldots + a_k x^k$ as a coefficient vector $a = (a_0, a_1, \ldots, a_k) \in \mathbb{R}^k$, e.g. $x$ is represented as $(0, 1)$.
Then we can define an algebra among coefficient vectors, including a redefinition of additive identity and multiplicative identity.
To avoid potential confusion, let us denote the additive identity as $\mymathbb{0}$, and the multiplicative identity is $\mymathbb{1}$.
The algebra between the polynomials number $a$ of order $k_a$ and $b$ of order $k_b$ is defined as
\begin{equation}
    \eqname{PN}
    \begin{split}
    a \oplus b &= (a_0 + b_0, a_1 + b_1, \ldots, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_0 + b_0, a_1b_0 + a_0b_1, a_{2}b_{0} + a_{1}b_{1} + a_{0}b_{2},  \ldots, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (),  \\
    \mymathbb{1} &= (1), \label{eq:polynomial}
    \end{split}
\end{equation}
where $\oplus$ and $\odot$ are the standard polynomial addition and multiplication operations.

@@

The configuration enumeration of independent sets include, for example, the enumeration of all independent sets, the enumeration of all MISs, and the enumeration of independent sets with the largest several sizes.
Recall that in the definition of a vertex tensor in \Cref{eq:vertextensor}, variables carry labels, so that one can read out all independent sets directly from the output polynomials.
The multiplication between labeled variables is commutative while the summation of labeled variables forms a set.
Intuitively, one can use a bit string as the representation of a labeled variable and use the bit-wise or $\lor^\circ$ as the multiplication operation.
For example, in a 5-vertex graph, $x_2$ and $x_5$ can be represented as $01000$ and $00001$ respectively and their multiplication $x_2x_5$ can be represented as $01000 \lor^\circ 00001 = 01001$.
To enumerate all independent sets, we designed an algebra on sets of bit strings:
\begin{equation}
\eqname{SN}
\begin{split}
    s \oplus t &= s \cup t\\
    s \odot t &= \{\sigma \lor^\circ \tau \, \mid \, \sigma \in s, \tau \in t\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{0^{\otimes |V|}\},
\end{split}\label{eq:set}
\end{equation}
where $s$ and $t$ are each a set of $|V|$-bit strings.
\begin{example}\label{eg:setalgebra}
    For elements being bit strings of length $5$, we have the following set algebra
\begin{equation*}
\begin{split}
    &\{00001\} \oplus \{01110, 01000\} = \{01110, 01000\} \oplus \{00001\} = \{00001,01110, 01000\}\\
    &\{00001\} \oplus \{\} = \{00001\}\\
&\\
    &\{00001\} \odot \{01110, 01000\} = \{01110, 01000\} \odot \{00001\} = \{01111, 01001\}\\
    &\{00001\} \odot \{\} = \{\}\\
    &\{00001\} \odot \{00000\} = \{00001\}.
\end{split}
\end{equation*}
\end{example}

@@

\section{Hard problems and tensor networks}\label{sec:otherproblems}
\subsection{Maximal independent sets and maximal cliques}\label{sec:maximal}
In this section, we focus the discussion on the maximal independent sets problem since finding maximal cliques of a graph is equivalent to finding the maximal independent sets of its complement.
Let $G=(V,E)$ be a graph, we denote the neighborhood of a vertex $v\in V$ as $N(v)$.
A maximal independent set $I_m$ is an independent set such that no $v \in V$ satisfies $I_m \cap (\{v\} \cup N[v])  = \emptyset$, i.e.\ an independent set that cannot become a larger one by adding a new vertex.
The corresponding tensor network $\mathcal{N}_{\rm mIS}(G)$ can be defined as
\begin{equation}\label{eq:maxistensornetwork}
\begin{split}
    \Sigma &= \{s_v \mid v \in V\},\\
    \mathcal{T} &= \{T^{(v)}_{s_{N(v,1)} s_{N(v, 2)}\ldots s_{N(v,d(v))}s_v} \mid v\in V\},\\
    O &= P,\\
\end{split}
\end{equation}
where we defined a tensor for each $v \in V$ and its neighborhood $N(v)$ as
\begin{equation}\label{eq:maximal}
    T^{(v)}_{s_{N(v, 1)}s_{N(v, 2)}\ldots s_{N(v, d(v))}s_v} = \begin{cases}
        s_vx_v^{w_v} & s_{N(v, 1)}=s_{N(v, 2)}=\ldots=s_{N(v, d(v))}=0,\\
        1-s_v& \text{otherwise}.
    \end{cases}
\end{equation}
Here, $N(v, k)$ is the $k$th vertex in $N(v)$ and $d(v) = |N(v)|$ is the degree of vertex $v$.
If $s_v = 1$, then none of its neighbours can be a member of $I_{m}$ by the independence constraint, contributing a factor $x_v^{w_v}$.
If $s_v = 0$, then at least on of its neighbors must be in $I_{m}$ by the maximal constraint, contributing a unit factor.
For a degree 2 vertex $v$, the tensor has the following form
\begin{equation}
    T^{(v)}=\left(\begin{matrix}
    \left(\begin{matrix}
        ~~0 &~1 \\
        ~~1 &~1
    \end{matrix}\right)\\
    \left(\begin{matrix}
        x_v^{w_v} &0 \\
        0 &0
    \end{matrix}\right)
    \end{matrix}\right).
\end{equation}
\begin{theorem}\label{thm:miscomplex}
    The tensor network representation of a maximal independent set problem on graph $G=(V,E)$ (\Cref{eq:maxistensornetwork}) can be contracted in $cc(\mathcal{N}_{mIS}(G)) = O(|V|)2^{O({\rm tw}(G)\Delta)}$ number of additions and multiplications, where $\Delta$ is the maximum degree of vertices in $G$.
\end{theorem}
\begin{proof}
    The prefactor $|V|$ comes from the number of tensors,
    while the contraction complexity of pair-wise tensor contraction is closely related to the treewidth of the line graph of its hypergraph representation ${\rm tw}(L(\mathcal{N}_{mIS}(G)))$.
    In the following, we will show this quantity is upper bounded by $(\Delta+1)$ times the treewidth of $G$.
    In the line graph $L(\mathcal{N}_{mIS}(G))$, a tensor $T^{(v)}$ corresponds to a clique over $N(v) \cup \{v\}$.
    In the following, we will show given an optimal tree decomposition of $G$, it is always possible to include all cliques into the bags by increasing the bag size by a factor $(\Delta + 1)$.
    Let $v \in V$ be a vertex and $N(v)$ be its neighborhood, we first arbitrarily pick an edge $u \in N(v)$,
    then by the definition of tree decomposition, we can find a bag containing this edge, and lastly we include all $N(v) \cup \{v\}$ into this bag.
    Hence the maximum tensor rank during contraction is upper bounded by ${\rm tw}(G)\Delta $, proving the theorem.
\end{proof}
Let us consider the graph in \Cref{eg:tensorcontraction}. The corresponding tensor network structure for computing the maximal independent polynomial has the following hypergraph representation. \\

>>>>

\section{Tensor networks}\label{sec:tn}
A tensor network is a multi-linear map from a collection of labelled tensors $\mathcal{T}$ to an output tensor.
It is formally defined as follows.
\begin{definition}[Tensor Network~\cite{Cirac2021, Orus2014}]
    A tensor network a multi-linear map specified by a triple of $\mathcal{N} = (\Sigma, \Tensors, \boldsymbol{\sigma}_o)$,
    where $\Sigma$ is an alphabet of labels,
    $\Tensors = \{T^{(1)}_{\boldsymbol{\sigma}_1}, T^{(2)}_{\boldsymbol{\sigma}_2}, \ldots, T^{(M)}_{\boldsymbol{\sigma}_M}\}$ is a set of tensors as the inputs,
    and $\boldsymbol{\sigma}_o$ is a string of labels for the output tensor.
    Each $T^{(k)}_{\boldsymbol{\sigma_k}} \in \mathcal{T}$ is labelled by a string of labels $\boldsymbol{\sigma}_k \in \Sigma^{r(T^{(k)})}$, where $r(T^{(k)})$ the rank of $T^{(k)}$.
    The multi-linear map or the \emph{contraction} on this triple is
    \begin{equation}
        O_{\boldsymbol{\sigma}_o} = \sum_{\Sigma \setminus \sigma_o} \prod_{k=1}^{M} T^{(k)}_{\boldsymbol{\sigma_k}},
    \end{equation}
    where the summation runs over all possible configurations over the set of labels absent in the output tensor.
\end{definition}
%The operation to evaluate a tensor network is called \textit{contraction}, which is defined as a summation of tensor element products over the labels not appearing in the output tensor.
For example, the matrix multiplication can be specified as a tensor network
\begin{equation}
\mathcal{N}_{\rm matmul} = \left(\{i,j,k\}, \{A_{ij}, B_{jk}\}, ik\right),
\end{equation}
where $A_{ij}$ and $B_{jk}$ are input matrices (two-dimensional tensors), and $(i,k)$ are labels associated to the output.
The contraction is defined as $O_{ik} = \sum_j A_{ij}B_{jk}$, where the subscripts are for tensor indexing, and the tensor dimensions with a same label must have the same size.
%can be viewed as a generalization of binary matrix multiplication to nary tensor contraction.
%Generalized Einstein's notation is often used to represent a tensor network, where we use a label in the subscript to represent a degree of freedom.
%One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% Einstein's notation is often used to represent a tensor network, e.g.\ it represents the matrix multiplication between two matrices $A$ and $B$ as $C_{ik} = A_{ij}B_{jk}$,
% where we use a label in the subscript to represent a degree of freedom.
% One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice, either both in input tensors (which will be summed over) or one in a input tensor and another in the output tensor.
% Hence a tensor network can be represented as a open simple graph,
% where a tensor is mapped to a vertex, a label shared by two input tensors is mapped to a normal edge and a label appears in the output tensor is mapped to an open edge.
% In this work, we adopt a more general form of tensor networks that do not restrict how many times a label can appear in tensors. 
The graphical representation of a tensor network is an open (meaning an edge can connect to external vertices) hypergraph, where an input tensor is mapped to a vertex and a label is mapped to a hyperedge that can connect an arbitrary number of vertices, while the labels appearing in the output tensor are open.
Our notation is a minor generalization of the standard tensor network notation used in physics as we do not restrict the number of times a label can appear in the tensors. 
While this generalized form is equivalent in representation power, it can have smaller contraction complexity as will be illustrated in \Cref{sec:tensorbad}.
% because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors, a high dimensional equivalence of the identity matrix.
%However, introducing $\delta$ tensors may significantly increase the complexity of the tensor contraction. We illustrate this subtle point in \Cref{sec:tensorbad}.

\begin{example}
\begin{equation}
\begin{split}
    \Sigma &= \{i,j,k,l,m\},\\
    \mathcal{T} &= \{A_{jkm}, B_{mil}, V_{jm}\},\\
    \boldsymbol{\sigma}_o &= ijk,\\
    \end{split}
\end{equation}
is a tensor network that can be evaluated as $O_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$.
Its hypergraph representation is shown below, where we use different colors to represent different hyperedges.

@@

\section{Tensor network representation of independent sets} \label{sec:tnmap}
This section describes the reduction of the independent set problem to a tensor network contraction problem.
An alternative interpretation, perhaps more accessible to physicists, can be found in \Cref{sec:energymodel}, where we introduce the reduction from the energy model of hardcore lattice gases~\cite{Dyre2016, Fernandes2007}.
Let $G=(V,E)$ be an undirected graph with each vertex $v\in V$ being associated with a weight $w_v$.
An independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, we have $(u, v) \not\in E$.
To the independent set problem on $G$ can be encoded as a tensor network $\mathcal{N}_{\rm IS}(G)$
\begin{equation}\label{eq:mistensornetwork}
\begin{split}
    \Sigma &= \{s_v \mid v \in V\},\\
    \mathcal{T} &= \{W^{(v)}_{s_v} \mid v\in V\} \cup \{B^{(u, v)}_{s_us_v} \mid (u, v) \in E\},\\
    \boldsymbol{\sigma}_o &= \varepsilon,
    \end{split}
\end{equation}
where for each vertex $v$, we define a parameterized rank-one tensor associated with it as
\begin{equation}
    W^{(v)} = \left(\begin{matrix}
        1 \\
        x_v^{w_v}
    \end{matrix}\right),\label{eq:vertextensor}
\end{equation}
and for each edge $(u, v) \in E$, we define a matrix $B$ indexed by $s_u$ and $s_v$ as
\begin{equation}
    \qquad \quad 
       B^{(u, v)} = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right). \label{eq:edgetensor}
\end{equation}
We map each vertex $v\in V$ to a label $s_v \in \{0, 1\}$ of dimension $2$, where we use $0$ ($1$) to denote a vertex is absent (present) in $I$.
%We use the superscripts to denote which vertex or edge a tensor is associated to.
These labels can be used as subscripts of tensors to index tensor elements, e.g.\ $W^{(v)}_0=1$ is the first element associated with $s_v=0$ and $W^{(v)}_1=x_v^{w_v}$ is the second element associated with $s_v=1$, where $x_v$ is a variable associated with vertex $v$.
The labels associated to the output tensor is an empty string $\varepsilon$, meaning this tensor has rank $0$, i.e. a scalar.
The independence constraint is encoded in the edge tensors, where we use $B_{11} = 0$ to denote two vertices connected by an edge cannot be both in the independent set. 
The contraction of this tensor network is
\begin{equation}\label{eq:idp}
    P(G) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{v\in V} W^{(v)}_{s_v} \prod\limits_{(u,v) \in E} B^{(u,v)}_{s_u s_v},
\end{equation}
where the summation runs over all $2^{|V|}$ vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$. 
% It is easy to see that an edge tensor element $B_{s_us_v}$ encodes the independent set constraint. It zeros out the product whenever both $u$ and $v$ are in the independent set.
A vertex tensor element $W^{(v)}_{s_v}$ contributes a multiplicative factor $x_v^{w_v}$ whenever $v$ is in the set.

@@

\begin{example}\label{eg:tensorcontraction}
In this example, we map a 5-vertex graph (left) to a tensor network (right) and show how the contraction order reduces the time and space complexities.

@@

Let $x_i = x$ and $w_i = 1$, \Cref{eq:idp} corresponds to the independence polynomial:
\begin{equation}\label{eq:idpdef}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$, $\alpha(G) \equiv \max_{I}\sum_{v\in I}w_v$ is the size of a maximum independent set and it is also called the independence number.
An independence polynomial is a useful graph characteristic related to, for example, the partition functions~\cite{Lee1952,Yang1952} and Euler characteristics of the independence complex~\cite{Bousquet2008, Levit2009}.
%Its connection to \Cref{eq:idp} can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
%and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. The summation counts the number of independent sets of size $k$. 
By assigning a real number to $x$, one can evaluate this independence polynomial for this specific value directly using tensor network contraction.
For example, the total number of independent sets can be evaluated as $I(G, 1)$.
However, instead of evaluating this polynomial for a certain value, we are more interested in knowing the coefficients of this polynomial, because this quantity tells us the counting of independent sets at different sizes.
%first elevate the tensor elements $0$s and $1$s in tensors $W(x)$ and $B$ from the standard number types to the additive identity, $\mymathbb{0}$, and multiplicative identity, $\mymathbb{1}$, of a commutative semiring as discussed in \Cref{sec:generic}.
%\purple{I'd
%say that you're about to describe this semiring, otherwise it seems like you're going to
%use the general definition of the semiring rather than a specific type of semiring}\blue{from $1$ to $\mymathbb{1}$ and $0$ to $\mymathbb{0}$ applies for a general semiring, I haven't change this part, let's discuss to make sure I understand your point.}\purple{Perhaps start the sentence with: Our semiring elements are a polynomial type...}
To this end, let us create a polynomial number data type by representing a polynomial $a_0 + a_1 x + \ldots + a_k x^k$ as a coefficient vector $a = (a_0, a_1, \ldots, a_k) \in \mathbb{R}^k$, e.g. $x$ is represented as $(0, 1)$.
Then we can define an algebra among coefficient vectors, including a redefinition of additive identity and multiplicative identity.
To avoid potential confusion, let us denote the additive identity as $\mymathbb{0}$, and the multiplicative identity is $\mymathbb{1}$.
The algebra between the polynomials number $a$ of order $k_a$ and $b$ of order $k_b$ is specified as
\begin{equation}
    \eqname{PN}
    \begin{split}
    a \oplus b &= (a_0 + b_0, a_1 + b_1, \ldots, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_0 + b_0, a_1b_0 + a_0b_1, a_{2}b_{0} + a_{1}b_{1} + a_{0}b_{2},  \ldots, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (),  \\
    \mymathbb{1} &= (1), \label{eq:polynomial}
    \end{split}
\end{equation}
where $\oplus$ and $\odot$ are the standard polynomial addition and multiplication operations.

@@

The configuration enumeration of independent sets include, for example, the enumeration of all independent sets, the enumeration of all MISs, and the enumeration of independent sets with the largest several sizes.
Recall that in the definition of a vertex tensor in \Cref{eq:vertextensor}, variables carry labels, so that one can read out all independent sets directly from the output polynomials.
The multiplication between labelled variables is commutative while the summation of labelled variables forms a set.
Intuitively, one can use a bit string as the representation of a labelled variable and use the bit-wise or $\lor^\circ$ as the multiplication operation.
For example, in a 5-vertex graph, $x_2$ and $x_5$ can be represented as $01000$ and $00001$ respectively and their multiplication $x_2x_5$ can be represented as $01000 \lor^\circ 00001 = 01001$.
To enumerate all independent sets, we designed an algebra on sets of bit strings:
\begin{equation}
\eqname{SN}
\begin{split}
    s \oplus t &= s \cup t\\
    s \odot t &= \{\sigma \lor^\circ \tau \, \mid \, \sigma \in s, \tau \in t\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{0^{\otimes |V|}\},
\end{split}\label{eq:set}
\end{equation}
where $s$ and $t$ are each a set of $|V|$-bit strings.
\begin{example}\label{eg:setalgebra}
    For elements being bit strings of length $5$, we have the following set algebra
\begin{equation*}
\begin{split}
    &\{00001\} \oplus \{01110, 01000\} = \{01110, 01000\} \oplus \{00001\} = \{00001,01110, 01000\}\\
    &\{00001\} \oplus \{\} = \{00001\}\\
&\\
    &\{00001\} \odot \{01110, 01000\} = \{01110, 01000\} \odot \{00001\} = \{01111, 01001\}\\
    &\{00001\} \odot \{\} = \{\}\\
    &\{00001\} \odot \{00000\} = \{00001\}.
\end{split}
\end{equation*}
\end{example}

@@

\section{Hard problems and tensor networks}\label{sec:otherproblems}
\subsection{Maximal independent sets and maximal cliques}\label{sec:maximal}
In this section, we focus the discussion on the maximal independent sets problem since finding maximal cliques of a graph is equivalent to finding the maximal independent sets of its complement.
Let $G=(V,E)$ be a graph, we denote the neighborhood of a vertex $v\in V$ as $N(v)$.
A maximal independent set $I_m$ is an independent set such that no $v \in V$ satisfies $I_m \cap (\{v\} \cup N[v])  = \emptyset$, i.e.\ an independent set that cannot become a larger one by adding a new vertex.
The corresponding tensor network $\mathcal{N}_{\rm mIS}(G)$ can be specified as
\begin{equation}\label{eq:maxistensornetwork}
\begin{split}
    \Sigma &= \{s_v \mid v \in V\},\\
    \mathcal{T} &= \{T^{(v)}_{s_{N(v,1)} s_{N(v, 2)}\ldots s_{N(v,d(v))}s_v} \mid v\in V\},\\
    \boldsymbol{\sigma}_o &= \varepsilon,\\
\end{split}
\end{equation}
where we defined a tensor for each $v \in V$ and its neighborhood $N(v)$ as
\begin{equation}\label{eq:maximal}
    T^{(v)}_{s_{N(v, 1)}s_{N(v, 2)}\ldots s_{N(v, d(v))}s_v} = \begin{cases}
        s_vx_v^{w_v} & s_{N(v, 1)}=s_{N(v, 2)}=\ldots=s_{N(v, d(v))}=0,\\
        1-s_v& \text{otherwise}.
    \end{cases}
\end{equation}
Here, $N(v, k)$ is the $k$th vertex in $N(v)$ and $d(v) = |N(v)|$ is the degree of vertex $v$.
If $s_v = 1$, then none of its neighbours can be a member of $I_{m}$ by the independence constraint, contributing a factor $x_v^{w_v}$.
If $s_v = 0$, then at least on of its neighbors must be in $I_{m}$ by the maximal constraint, contributing a unit factor.
For a degree 2 vertex $v$, the tensor has the following form
\begin{equation}
    T^{(v)}=\left(\begin{matrix}
    \left(\begin{matrix}
        ~~0 &~1 \\
        ~~1 &~1
    \end{matrix}\right)\\
    \left(\begin{matrix}
        x_v^{w_v} &0 \\
        0 &0
    \end{matrix}\right)
    \end{matrix}\right).
\end{equation}
\begin{theorem}\label{thm:miscomplex}
    The tensor network representation of a maximal independent set problem on graph $G=(V,E)$ (\Cref{eq:maxistensornetwork}) can be contracted in $cc(\mathcal{N}_{mIS}(G)) = O(|V|)2^{O({\rm tw}(G)\Delta)}$ number of additions and multiplications, where $\Delta$ is the maximum degree of vertices in $G$.
\end{theorem}
\begin{proof}
    The prefactor $|V|$ comes from the number of tensors,
    while the contraction complexity of pair-wise tensor contraction is closely related to the treewidth of the line graph of its hypergraph representation ${\rm tw}(L(\mathcal{N}_{mIS}(G)))$.
    In the following, we will show this quantity is upper bounded by $(\Delta+1)$ times the treewidth of $G$.
    In the line graph $L(\mathcal{N}_{mIS}(G))$, a tensor $T^{(v)}$ corresponds to a clique over $N(v) \cup \{v\}$.
    In the following, we will show given an optimal tree decomposition of $G$, it is always possible to include all cliques into the bags by increasing the bag size by a factor $(\Delta + 1)$.
    Let $v \in V$ be a vertex and $N(v)$ be its neighborhood, we first arbitrarily pick an edge $u \in N(v)$,
    then by the definition of tree decomposition, we can find a bag containing this edge, and lastly we include all $N(v) \cup \{v\}$ into this bag.
    Hence the maximum tensor rank during contraction is upper bounded by ${\rm tw}(G)\Delta $, proving the theorem.
\end{proof}
Let us consider the graph in \Cref{eg:tensorcontraction}. The corresponding tensor network structure for computing the maximal independent polynomial has the following hypergraph representation. \\
